Supplementary Material for ``Evaluating infectious disease forecasts with allocation scoring rules''

### 2023-12-12 alloscore_manuscript/supplement.rnw
attempting to put stuff about dfq to rest

In order for a scoring rule to be guaranteed to yield logically consistent forecast evaluations, it must be \emph{proper}, which means,
in essence, that a forecaster cannot perceive any forecast at odds with their beliefs about the events being forecasted as having an optimal expected score.
We formalize this in terms of forecaster ``self-assessment'' in section \ref{sec:proper_def} and then address the propriety of the allocation score in three stages:
\begin{itemize}
  \item Section \ref{sec:alloscore_proper} shows that the allocation score is proper when the forecaster's actual forecast distribution $F$ is available (as is any score derived using the decision theoretic procedures outlined in section 2 of the manuscript)
  \item Section \ref{sec:distfromq_alloscore_proper_prospective} shows that the allocation score is still proper if the forecast is represented as a member of a parametric family of distributions, so long as this representation and scoring procedure was specified prospectively by the hub. This includes the use of methods such as \verb`distfromq` for constructing forecast distributions as a special case.
  \item Section \ref{sec:distfromq_alloscore_improper_post_hoc} shows that \emph{post hoc} application of this scoring procedure to quantile forecasts is improper, and makes some suggestions for organized forecasting efforts who are considering using the allocation score.
\end{itemize}

@@

In practice, open forecasting exercises are generally not able to collect an exact representation of the forecast distribution $F$ other than in simple settings such as for a categorical variable with a relatively small number of categories. In settings where the outcome being forecasted is a continuous quantity (such as rates of influenza-like illness among outpatient doctor visits) or a count (such as influenza hospitalizations), forecasting exercises have therefore resorted to collecting summaries of a forecast distribution such as bin probabilities or predictive quantiles. This raises the question of whether the allocation score is still proper if the forecast distribution $F$ is not itself directly recorded. The answer we provide in this section is that it is, so long as the fact that the allocation score will be used for forecast evaluation and the method that will be used to obtain a distributional estimate of $F$ from the provided forecast representation are communicated to participating forecasters prospectively.

We consider a setting where a forecasting exercise (such as a forecast hub) pre-specifies that forecasts will be represented using a parametric family of forecast distributions $F_\theta(y)$, and the task of the forecaster is to select a particular parameter value $\theta$. We use $\mathcal{P}$ to denote the collection of all distributions $F_\theta$ in the given parametric family. For instance, it has recently been proposed that mixture distributions could be used to represent forecast distributions \cite{wadsworth2023mixture}. Additionally, we note that the functionality in \verb`distfromq` can be viewed as specifying a parametric family $\mathcal{P}_{\mathrm{dfq}}$ where the parameters $\theta$ of $F_\theta$
are its quantiles at pre-specified probability levels, and where the shape of any $F_\theta \in \mathcal{P}_{\mathrm{dfq}}$ over the full range of its support is entirely controlled by these quantiles.

We find it helpful now to formally distinguish between two decision making problems. The first is the public health decision maker's allocation problem where the task is to select an allocation $x$, with the allocation loss $s_A(x, y) = \sum_{i=1}^N L \cdot \max(0, y_i - x_i)$ as described in section \ref{sec:methods.detailed.specific_allocation} of the main text.
The second is the forecaster's reporting problem where the task is to select parameter values $\theta$ to report. The forecaster's loss is given by $s_R(\theta, y) = s_A(x^{F_\theta}, y)$, where $x^{F_\theta}$ is the Bayes act for the allocation problem under the distribution $F_\theta$. In words, the loss associated with reporting $\theta$ is equal to the loss associated with taking the Bayes allocation corresponding to the distribution $F_\theta$.

Following our usual construction, the Bayes act for the forecast reporting problem is the parameter set that minimizes the forecaster's expected loss:
\begin{align*}
\theta^F &= \text{argmin}_\theta \Ex_F [s_R(\theta, Y)] \\
&= \text{argmin}_\theta \Ex_F [s_A(x^{F_\theta}, Y)]
\end{align*}
This leads the forecaster to choose values of the parameters $\theta$ such that the allocation $x^{F_\theta}$ is as close as possible to the allocation $x^F$ under the forecaster's true forecast distribution $F$.
\apg{I think we need to say something about what we mean by ``close''.  The score divergence is what we are confronting here and I appreciate that 
others may not want to broach the topic.  But I think we have to say something about ``score expectation defined distance''. This also relates directly to the notation crunch below.}

We then arrive at the scoring rule
$$S_R(F, y) = s_R(\theta^F, y) = s_A(x^{F_{\theta^F}}, y).$$
\elr{Notation got pretty bad here... borderline about whether it's worth searching for a better notation for "the bayes allocation for the distribution with the bayes parameters"}
It follows from the discussion in section \ref{sec:alloscore_proper} that this is a proper scoring rule for $F$. Additionally, we note that the score can be calculated from the reported parameter values $\theta^F$.

We emphasize that the forecaster's true predictive distribution $F$ does not need to be a member of the specified parametric family $\mathcal{P}$ for this construction to yield a proper score.
It is, however, necessary to specify the parametric family to use and the foundational scoring rule $s_A$ (including any relevant problem parameters such as the resource constraint $K$) in advance, so that forecasters can identify the Bayes act parameter set $theta^F$ to report.

As an alternative, a forecast hub could ask forecasters to provide the Bayes allocations associated with their forecasts for one or more fixed resource constraints $K$, allowing for direct calculation of scores and narrowing the gap between model outputs and public health end users.

\subsection{Post hoc evaluation of quantile forecasts using distfromq and the allocation score is improper}
\label{sec:distfromq_alloscore_improper_post_hoc}

In this section, we illustrate that a post hoc evaluation of quantile forecasts using \verb`distfromq` and the allocation score is not proper. In brief, this is because the Bayes parameters $\theta^F$ for the forecaster to report may not align with quantiles of the distribution $F$ at the specified probability levels. We illustrate this with examples (borrow from elicitation.R...)

>>>>

In order for a scoring rule to be guaranteed to yield logically consistent forecast evaluations, it must be \emph{proper}, which means,
in essence, that a forecaster cannot perceive any forecast that is at odds with their beliefs about the events being forecasted as having an optimal expected score.
We formalize this in terms of forecaster ``self-assessment'' in section \ref{sec:proper_def} and then address the propriety of the allocation score in two stages:
\begin{itemize}
  \item Section \ref{sec:alloscore_proper} shows that the allocation score is proper when the forecaster's actual forecast distribution $F$ is available (as is any score derived using the decision theoretic procedures outlined in section 2 of the manuscript)
  \item Section \ref{sec:distfromq_alloscore_proper_prospective} addresses the propriety of an analysis that combines an approximate parametric representation of the forecast distribution and evaluation with the allocation score. It shows that this procedure yields a proper score so long as this representation and scoring procedure was specified prospectively by the hub. This includes the use of methods such as \verb`distfromq` for constructing forecast distributions as a special case. However,  \emph{post hoc} application of this scoring procedure to quantile forecasts is improper.
\end{itemize}

@@

In practice, open forecasting exercises are generally not able to collect an exact representation of the forecast distribution $F$ other than in simple settings such as for a categorical variable with a relatively small number of categories. In settings where the outcome being forecasted is a continuous quantity (such as rates of influenza-like illness among outpatient doctor visits) or a count (such as influenza hospitalizations), forecasting exercises have therefore resorted to collecting summaries of a forecast distribution such as bin probabilities or predictive quantiles. This raises the question of whether use of the allocation score still consitutes a proper evaluation procedure if the forecast distribution $F$ is not itself directly recorded. The answer we provide in this section is that it is, so long as the fact that the allocation score will be used for forecast evaluation and the method that will be used to obtain a distributional estimate of $F$ from the provided forecast representation are communicated to participating forecasters prospectively.

We consider a setting where a forecasting exercise (such as a forecast hub) pre-specifies that forecasts will be represented using a parametric family of forecast distributions $G_\theta(y)$, and the task of the forecaster is to select a particular parameter value $\theta$. We use $\mathcal{P}$ to denote the collection of all distributions $G_\theta$ in the given parametric family. For instance, it has recently been proposed that mixture distributions could be used to represent forecast distributions \cite{wadsworth2023mixture}. Additionally, we note that the functionality in \verb`distfromq` can be viewed as specifying a parametric family $\mathcal{P}_{\mathrm{dfq}}$ where the parameters $\theta$ of $G_\theta$
are its quantiles at pre-specified probability levels, and where the shape of any $G_\theta \in \mathcal{P}_{\mathrm{dfq}}$ over the full range of its support is entirely controlled by these quantiles.

We find it helpful now to formally distinguish between two decision making problems. The first is the public health decision maker's allocation problem where the task is to select an allocation $x$, with the allocation loss $s_A(x, y) = \sum_{i=1}^N L \cdot \max(0, y_i - x_i)$ as described in section \ref{sec:methods.detailed.specific_allocation} of the main text.
The second is the forecaster's reporting problem where the task is to select parameter values $\theta$ to report. The forecaster's loss is given by $s_R(\theta, y) = s_A(x^{G_\theta}, y)$, where $x^{G_\theta}$ is the Bayes act for the allocation problem under the distribution $G_\theta$. In words, the loss associated with reporting $\theta$ is equal to the loss associated with taking the Bayes allocation corresponding to the distribution $G_\theta$.

Following our usual construction, the Bayes act for the forecast reporting problem is the parameter set that minimizes the forecaster's expected loss. Breaking with our earlier notation for improved legibility, we use $\theta^\star(F)$ to denote this Bayes act:
\begin{align*}
\theta^\star(F) &= \text{argmin}_\theta \Ex_F [s_R(\theta, Y)] \\
&= \text{argmin}_\theta \Ex_F [s_A(x^{G_\theta}, Y)]
\end{align*}

We then arrive at the scoring rule
$$S_R(F, y) = s_R(\theta^\star(F), y) = s_A(x^{G_{\theta^\star(F)}}, y).$$
It follows from the discussion in section \ref{sec:alloscore_proper} that this is a proper scoring rule for $F$. Additionally, we note that the score can be calculated from the reported parameter values $\theta^\star(F)$.

We emphasize that the forecaster's true predictive distribution $F$ does not need to be a member of the specified parametric family $\mathcal{P}$ for this construction to yield a proper score.
It is, however, necessary to specify the parametric family to use and the foundational scoring rule $s_A$ (including any relevant problem parameters such as the resource constraint $K$) in advance, so that forecasters can identify the Bayes act parameter set $\theta^\star(F)$ to report.

For a given resource constraint $K$, if it is possible to find a member of the specified parametric family $\mathcal{P}$ with the same allocation as $F$ for all possible forecast distributions $F$ (i.e., for any $F$ there is a $G_\theta \in \mathcal{P}$ such that $x^F = x^{G_\theta}$), the scoring rule $S_R$ will yield the same score as the allocation score $S_A$.
For the particular choice of the parametric family $\mathcal{P}_{dfq}$ (i.e., using the \verb`distfromq` package), this is the case. For instance, the forecaster could pick one required quantile level (such as 0.5, for which the corresponding predictions are predictive medians), and set the submitted quantiles of their forecast distribution at that level to be the desired allocations.
However, this representation of the forecast may be quite different from the actual forecast distribution $F$.

A key observation is that a \emph{post hoc} analysis combining \verb`distfromq` with the allocation score does not in general yield a proper evaluation method. This is because the forecast distribution $F$ and the distribution $G_\theta \in \mathcal{P}_{dfq}$ with the same quantiles as $F$ may determine quite different resource allocations, particularly if the tail extrapolations performed by \verb`distfromq` do not match the tail behavior of $F$.

As another alternative for practical forecasting exercises, a forecast hub could ask forecasters to provide the Bayes allocations associated with their forecasts for one or more fixed resource constraints $K$. This would allow for direct calculation of scores and narrowing the gap between model outputs and public health end users. For this to be feasible, implementations of the allocation algorithm would have to be provided to participating forecasters.
